---
title: "POMPE MDS plot"
author: "JCMM"
date: "8/3/2021"
output: pdf_document
---

## Install and load packages
[Elegant-way-to-check-for-missing-packages-and-install-them](https://stackoverflow.com/questions/4090169/elegant-way-to-check-for-missing-packages-and-install-them)

```{r, install and load packages}
list_of_packages <- c("dplyr", "grafify", "RColorBrewer", "ggpubr")

install.packages.auto <- function(x) { 
  if(isTRUE(x %in% .packages(all.available = TRUE))) { 
    eval(parse(text = sprintf("require(\"%s\")", x)))
  } else { 
    #update.packages(ask= FALSE) #update installed packages.
    eval(parse(text = sprintf("install.packages(\"%s\", dependencies = TRUE)", x)))
  }
  if(isTRUE(x %in% .packages(all.available=TRUE))) { 
    eval(parse(text = sprintf("require(\"%s\")", x)))
  } else {
    if (!requireNamespace("BiocManager", quietly = TRUE))
      install.packages("BiocManager")
    eval(parse(text = sprintf("BiocManager::install(\"%s\")", x, update = FALSE)))
    eval(parse(text = sprintf("require(\"%s\")", x)))
  }
}

lapply(list_of_packages, function(x) {message(x); install.packages.auto(x)})

rm(list_of_packages, install.packages.auto)

```

```{r }
# Change global default setting so every data frame created will not auto-convert to factors unless explicitly instructed
options(stringsAsFactors = FALSE)
```

## Load 1000 G ped file
```{r, load 1000 G metadata}
PED <- read.table('../Files/20130606_g1k.ped', 
                  header = TRUE, skip = 0, sep = '\t')
#reduce
PED = PED %>% select(Individual.ID, Population)
PED = rename(PED, IID = Individual.ID)
```

##Load Population code
from: [ftp://ftp.1000genomes.ebi.ac.uk/vol1/ftp/phase3/20131219.populations.tsv](ftp://ftp.1000genomes.ebi.ac.uk/vol1/ftp/phase3/20131219.populations.tsv)
```{r, Load Population code}

# 20131219.populations.tsv
POP <- read.table('../Files/20131219.populations.tsv', 
                 header = TRUE,  sep = '\t')

# reduce
POP = POP %>% select(Population.Code, Super.Population)
POP = rename(POP, Population = Population.Code)

#Annotate PED with groups from POP
PED_ann = merge(POP, PED, by = "Population")
```

##Load mds cluster result
```{r, load mds cluster result}

mds.cluster <- read.csv("../Files/1kG_POMPE_R.mds.csv", 
                       header=T)
G1000.cluster = merge(mds.cluster, PED_ann, by = "IID")

POMPE.cluster = mds.cluster[mds.cluster$FID!=0,]
#check number of POMPE patients, should be 130
dim(POMPE.cluster)

POMPE.cluster$Population= "Dutch Pompe"
POMPE.cluster$Super.Population= "Dutch Pompe"
POMPE.cluster = POMPE.cluster %>%  select(names(G1000.cluster))
#put them back together 
mds_df = rbind( G1000.cluster, POMPE.cluster)

```

## Plot axis C1 and C2 for all groups
```{r, plot}
#create a color pallette
super_pop_pal = c(
  "#984EA3",       #AFR
  "lightskyblue1", #AMR
  "#E41A1C",       #POMPE
  "#FF7F00",       #EAS
  "black",         #EUR
  "green"          #SAS
  ) 
  
Fig.a1 <- plot_xy_CatGroup(mds_df,
                          C2,
                          C1,
                          Super.Population,
                          symsize = 3,
                          symthick = 0.1,
                          fontsize = 15,
                          s_alpha = 0.5) +
  scale_fill_manual(values = super_pop_pal, name = "") +
  guides(fill = guide_legend(override.aes = list(alpha = 1)))  # override.aes overwrites the alpha

Fig.a1
```

## Plot axis C1 and C3 for all groups
```{r, plot}
#create a color pallette
super_pop_pal = c(
  "#984EA3",       #AFR
  "lightskyblue1", #AMR
  "#E41A1C",       #POMPE
  "#FF7F00",       #EAS
  "black",         #EUR
  "green"          #SAS
  ) 
  
Fig.a2 <- plot_xy_CatGroup(mds_df,
                          C3,
                          C1,
                          Super.Population,
                          symsize = 3,
                          symthick = 0.1,
                          fontsize = 15,
                          s_alpha = 0.5) +
  scale_fill_manual(values = super_pop_pal, name = "") +
  guides(fill = guide_legend(override.aes = list(alpha = 1)))  # override.aes overwrites the alpha

Fig.a2
```        

## Arrange panels A
```{r}

Fig.a <- ggarrange( Fig.a1 ,Fig.a2,
          common.legend = TRUE, legend = "right",
          labels = "A",  font.label = list(size = 18),
          ncol = 2, nrow = 1)
Fig.a

```


##Zoom in the European group
```{r}
#create a color pallette
EUR_pal = c(
  "blue",    #CEU
  "red" ,    #POMPE
  "#4DAF4A", #FIN*
  "gray48",  #GBR*
  "#FFFF33", #IBS*
  "#A65628"  #TSI*
  ) 

#remove non-EUR
EUR = POP %>% filter(Super.Population == "EUR") 
#plot
Fig.b1 <- plot_xy_CatGroup(mds_df %>% 
                            filter(Population == "Dutch Pompe" | Population %in% EUR$Population),
                          C2,
                          C1,
                          Population,
                          symsize = 3,
                          symthick = 0.1,
                          fontsize = 15,
                          s_alpha = 0.5) +
  # ylim(-0.04,0) +
  # xlim(-0.06,-0.035) +
  scale_fill_manual(values = EUR_pal, name = "") +
  guides(fill = guide_legend(override.aes = list(alpha = 1)))  # override.aes overwrites the alpha

Fig.b1
```



```{r}
#create a color pallette
EUR_pal = c(
  "blue",    #CEU
  "red" ,    #POMPE
  "#4DAF4A", #FIN*
  "gray48",  #GBR*
  "#FFFF33", #IBS*
  "#A65628"  #TSI*
  ) 

#remove non-EUR
EUR = POP %>% filter(Super.Population == "EUR") 

Fig.b2 <- plot_xy_CatGroup(mds_df %>% 
                            filter(Population == "Dutch Pompe" | Population %in% EUR$Population),
                          C3,
                          C1,
                          Population,
                          symsize = 3,
                          symthick = 0.1,
                          fontsize = 15,
                          s_alpha = 0.5) +
  # ylim(-0.04,0) +
  # xlim(-0.06,-0.035) +
  scale_fill_manual(values = EUR_pal, name = "") +
                    # 
  guides(fill = guide_legend(override.aes = list(alpha = 1)))  # override.aes overwrites the alpha

Fig.b2
```

## Arrange panels B
```{r}

Fig.b <- ggarrange( Fig.b1 ,Fig.b2,
          common.legend = TRUE, legend = "right",
          labels = "B",  font.label = list(size = 18),
          ncol = 2, nrow = 1)
Fig.b

```

## Arrange panels A and B
```{r}

Fig <- ggarrange( Fig.a ,Fig.b,
          ncol = 1, nrow = 2)
Fig
```

##If everyfing looks fine save figures as pdf
```{r}

ggsave("../Plots/POMPE_MDS_plot.AB.pdf", plot = Fig, dpi = 300,
       width = 30,  height = 20,
       units = "cm"
       )
ggsave("../Plots/POMPE_MDS_plot.A.pdf", plot = Fig.a, dpi = 300,
       width = 30,  height = 10,
       units = "cm"
       )
ggsave("../Plots/POMPE_MDS_plot.B.pdf", plot = Fig.b, dpi = 300,
       width = 30,  height = 10,
       units = "cm"
       )
```
